New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices

### 2024-09-16 jso.qmd
update biblio
doi and other bibliography fixes

  The projection pursuit (PP) guided tour interactively optimises a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections. Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optima has a small "squint angle", detectable only from close proximity. To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes. The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers. Additionally, methods for calculating the smoothness and squintability properties of the PP index are proposed. They are used to assess the optimiser performance in the presence of PP index complexities. A simulation study illustrates the use of these performance metrics to compare the JSO with existing optimisation methods available for the guided tour.  The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability are available in the `ferrn` package. 

@@

Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data. This is useful for exploring high-dimensional data, and creating plots of the data that reveal the main features to use for publication. The method involves optimising an objective function known as the PP index (e.g. @hall1989polynomial, @cook1993projection,
@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes interesting or informative projections. Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space. The index function $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function that measures an interesting aspect of the projected data, such as deviation from normality, presence of clusters, non-linear structure, or other features of interest. For a fixed sample of data, PP finds the orthonormal basis $A$ that maximises the index value of the projection, $Y = XA$:

@@

Optimisation of PP is often discussed when new indexes are proposed [@posse1995; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima. An implementation is available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105 illustrated how to diagnose optimisation processes, particularly focusing on the guided tour, and revealed a need for improved optimisation. While improving the quality of the optimisation solutions in the tour is essential, it is also important to be able to view the data projections as the optimisation progresses. Integrating the guided tour with a global optimisation algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process is a goal.

>>>>

  The projection pursuit (PP) guided tour interactively optimises a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections. Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optima has a small "squint angle", detectable only from close proximity. To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes. The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers. Additionally, methods for calculating the smoothness and squintability properties of the PP index are proposed. They are used to assess the optimiser performance in the presence of PP index complexities. A simulation study illustrates the use of these performance metrics to compare the JSO with existing optimisation methods available for the guided tour. The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability are available in the `ferrn` package. 

@@

Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data. This is useful for exploring high-dimensional data, and creating plots of the data that reveal the main features to use for publication. The method involves optimising an objective function known as the PP index (e.g. @hall1989polynomial, @cook1993projection,
@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criterion for what constitutes interesting or informative projections. Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space. The index function $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function that measures an interesting aspect of the projected data, such as deviation from normality, presence of clusters, non-linear structure, or other features of interest. For a fixed sample of data, PP finds the orthonormal basis $A$ that maximises the index value of the projection, $Y = XA$:

@@

Optimisation of PP is often discussed when new indexes are proposed [@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima. An implementation is available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105 illustrated how to diagnose optimisation processes, particularly focusing on the guided tour, and revealed a need for improved optimisation. While improving the quality of the optimisation solutions in the tour is essential, it is also important to be able to view the data projections as the optimisation progresses. Integrating the guided tour with a global optimisation algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process is a goal.

### 2024-09-16 jso.tex
update biblio
doi and other bibliography fixes

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@austin.utexas.edu}{\nolinkurl{huize.zhang@austin.utexas.edu}}. Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}. Nicolas
Langrené. Email: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}. Jessica
Wai Yin
Leung. Email: \href{mailto:Jessica.Leung@monash.edu}{\nolinkurl{Jessica.Leung@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Statistics and Data
Sciences, University of Texas at Austin, Austin, United
States\\ \textsuperscript{2} Department of Econometrics and Business
Statistics, Monash
University, Melbourne, Australia\\ \textsuperscript{3} Department of
Mathematical Sciences, Guangdong Provincial Key Laboratory of
Interdisciplinary Research and Application for Data Science,
BNU-HKBU~United~International~College, Zhuhai, China
\begin{abstract}
The projection pursuit (PP) guided tour interactively optimises a
criteria function known as the PP index, to explore high-dimensional
data by revealing interesting projections. Optimisation of some PP
indexes can be non-trivial, if they are non-smooth functions, or the
optima has a small ``squint angle'', detectable only from close
proximity. To address these challenges, this study investigates the
performance of a recently introduced swarm-based algorithm, Jellyfish
Search Optimiser (JSO), for optimising PP indexes. The performance of
JSO for visualising data is evaluated across various hyper-parameter
settings and compared with existing optimisers. Additionally, methods
for calculating the smoothness and squintability properties of the PP
index are proposed. They are used to assess the optimiser performance in
the presence of PP index complexities. A simulation study illustrates
the use of these performance metrics to compare the JSO with existing
optimisation methods available for the guided tour. The JSO algorithm
has been implemented in the R package, \texttt{tourr}, and functions to
calculate smoothness and squintability are available in the
\texttt{ferrn} package.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
projection pursuit\sep jellyfish search optimiser
(JSO)\sep optimisation\sep grand tour\sep high-dimensional data\sep 
exploratory data analysis
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi

@@

Projection pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974),
Huber (1985)) is a dimension reduction technique aimed at identifying
informative linear projections of data. This is useful for exploring
high-dimensional data, and creating plots of the data that reveal the
main features to use for publication. The method involves optimising an
objective function known as the PP index (e.g. Hall (1989), Cook, Buja,
and Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
(2020)), which defines the criteria for what constitutes interesting or
informative projections. Let \(X \in \mathbb{R}^{n\times p}\) be the
data matrix, \(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix,
where \(A\) belongs to the Stiefel manifold
\(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\) is a
linear transformation that maps data from a \(p\)-dimensional space into
a \(d\)-dimensional space. The index function
\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar function
that measures an interesting aspect of the projected data, such as
deviation from normality, presence of clusters, non-linear structure, or
other features of interest. For a fixed sample of data, PP finds the
orthonormal basis \(A\) that maximises the index value of the
projection, \(Y = XA\):

@@

Optimisation of PP is often discussed when new indexes are proposed
(Posse 1995a; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
index, when they introduced the PPGT, which monitors the optimisation
visually so that the user can see the projected data leading in and out
of the optima. An implementation is available in the \texttt{tourr}
package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
(2021) illustrated how to diagnose optimisation processes, particularly
focusing on the guided tour, and revealed a need for improved
optimisation. While improving the quality of the optimisation solutions
in the tour is essential, it is also important to be able to view the
data projections as the optimisation progresses. Integrating the guided
tour with a global optimisation algorithm that is efficient in finding
the global optimal and enables viewing of the projected data during the
exploration process is a goal.

@@

\begin{itemize}
\tightlist
\item
  a pseudo-derivative, that searches locally for the best direction,
  based on differencing the index values for very close projections.
\item
  a brute-force optimisation (CRS).
\item
  a modified brute force algorithm described in Posse (1995b).
\item
  an essentially simulated annealing (Bertsimas and Tsitsiklis 1993)
  where the search space is reduced as the optimisation.
\item
  a very localised search, to take tiny steps to get closer to the local
  maximum.
\end{itemize}

@@

\begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, colback=white, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, arc=.35mm, toprule=.15mm, rightrule=.15mm, left=2mm, bottomtitle=1mm]

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abdulah2023large}{}}%
Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
Genton, and Ying Sun. 2023. {``Large-Scale Environmental Data Science
with {ExaGeoStatR}.''} \emph{Environmetrics} 34 (1): e2770.
https://doi.org/\url{https://doi.org/10.1002/env.2770}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
Friedman, J. H., and J. W. Tukey. 1974. {``{A} {P}rojection {P}ursuit
{A}lgorithm for {E}xploratory {D}ata {A}nalysis.''} \emph{IEEE
Transactions on Computing C} 23: 881--89.

@@

\leavevmode\vadjust pre{\hypertarget{ref-grochowski2011}{}}%
Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
Based on Quality of Projected Clusters.''} In \emph{Adaptive and Natural
Computing Algorithms}, edited by Andrej Dobnikar, Uroš Lotrič, and
Branko Šter, 89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.

@@

\leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{Ann. Statist.} 13
(2): 435--75. \url{https://doi.org/10.1214/aos/1176349519}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-karvonen2023asymptotic}{}}%
Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
Estimates in {G}aussian Process Interpolation.''} \emph{SIAM/ASA Journal
on Uncertainty Quantification} 11 (4): 1225--57.

\leavevmode\vadjust pre{\hypertarget{ref-kr69}{}}%
Kruskal, J. B. 1969. {``Toward a Practical Method Which Helps Uncover
the Structure of a Set of Observations by Finding the Line
Transformation Which Optimizes a New {`Index of Condensation'}.''} In
\emph{Statistical Computation}, edited by R. C. Milton and J. A. Nelder,
427--40. New York: Academic Press.

@@

\leavevmode\vadjust pre{\hypertarget{ref-lee2010projection}{}}%
Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
Large p Small n Data.''} \emph{Statistics and Computing} 20 (3):
381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2018}{}}%
Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
Computational Approach.''} \emph{Computational Statistics and Data
Analysis} 120 (C): 42--57.
https://doi.org/\url{https://doi.org/10.1016/j.csda.2017.11.001}.

\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2020}{}}%
---------. 2020. {``Kurtosis-Based Projection Pursuit for Outlier
Detection in Financial Time Series.''} \emph{The European Journal of
Finance} 26 (2-3): 142--64.
https://doi.org/\url{https://doi.org/10.1080/1351847X.2019.1647864}.

\leavevmode\vadjust pre{\hypertarget{ref-marie-sainte2010}{}}%
Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
{``An Efficient Optimization Method for Revealing Local Optima of
Projection Pursuit Indices.''} In \emph{Swarm Intelligence}, edited by
Marco Dorigo, Mauro Birattari, Gianni A. Di Caro, René Doursat, Andries
P Engelbrecht, Dario Floreano, Luca Maria Gambardella, et al., 60--71.
Berlin, Heidelberg: Springer Berlin Heidelberg.

\leavevmode\vadjust pre{\hypertarget{ref-porcu2024matern}{}}%
Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
2024. {``The {M}atérn Model: A Journey Through Statistics, Numerical
Analysis and Machine Learning.''} \emph{Statistical Science} 39 (3):
469--92.

\leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
Posse, Christian. 1995b. {``Projection Pursuit Exploratory Data
Analysis.''} \emph{Computational Statistics and Data Analysis} 20 (6):
669--87.
https://doi.org/\url{https://doi.org/10.1016/0167-9473(95)00002-8}.

\leavevmode\vadjust pre{\hypertarget{ref-posse1995}{}}%
---------. 1995a. {``Projection Pursuit Exploratory Data Analysis.''}
\emph{Computational Statistics \& Data Analysis} 20 (6): 669--87.
https://doi.org/\url{https://doi.org/10.1016/0167-9473(95)00002-8}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tourr}{}}%
Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
Package for Exploring Multivariate Data with Projections.''}
\emph{Journal of Statistical Software} 40 (2): 1--18.
\url{http://doi.org/10.18637/jss.v040.i02}.

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@austin.utexas.edu}{\nolinkurl{huize.zhang@austin.utexas.edu}}. Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}. Nicolas
Langrené. Email: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}. Jessica
Wai Yin
Leung. Email: \href{mailto:Jessica.Leung@monash.edu}{\nolinkurl{Jessica.Leung@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Statistics and Data
Sciences, University of Texas at Austin, Austin, United
States\\ \textsuperscript{2} Department of Econometrics and Business
Statistics, Monash
University, Melbourne, Australia\\ \textsuperscript{3} Department of
Mathematical Sciences, Guangdong Provincial Key Laboratory of
Interdisciplinary Research and Application for Data Science,
BNU-HKBU~United~International~College, Zhuhai, China
\begin{abstract}
The projection pursuit (PP) guided tour interactively optimises a
criterion function known as the PP index, to explore high-dimensional
data by revealing interesting projections. Optimisation of some PP
indexes can be non-trivial, if they are non-smooth functions, or the
optima has a small ``squint angle'', detectable only from close
proximity. To address these challenges, this study investigates the
performance of a recently introduced swarm-based algorithm, Jellyfish
Search Optimiser (JSO), for optimising PP indexes. The performance of
JSO for visualising data is evaluated across various hyper-parameter
settings and compared with existing optimisers. Additionally, methods
for calculating the smoothness and squintability properties of the PP
index are proposed. They are used to assess the optimiser performance in
the presence of PP index complexities. A simulation study illustrates
the use of these performance metrics to compare the JSO with existing
optimisation methods available for the guided tour. The JSO algorithm
has been implemented in the R package, \texttt{tourr}, and functions to
calculate smoothness and squintability are available in the
\texttt{ferrn} package.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
projection pursuit\sep jellyfish search optimiser
(JSO)\sep optimisation\sep grand tour\sep high-dimensional data\sep 
exploratory data analysis
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi

@@

Projection pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974),
Huber (1985)) is a dimension reduction technique aimed at identifying
informative linear projections of data. This is useful for exploring
high-dimensional data, and creating plots of the data that reveal the
main features to use for publication. The method involves optimising an
objective function known as the PP index (e.g. Hall (1989), Cook, Buja,
and Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
(2020)), which defines the criterion for what constitutes interesting or
informative projections. Let \(X \in \mathbb{R}^{n\times p}\) be the
data matrix, \(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix,
where \(A\) belongs to the Stiefel manifold
\(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\) is a
linear transformation that maps data from a \(p\)-dimensional space into
a \(d\)-dimensional space. The index function
\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar function
that measures an interesting aspect of the projected data, such as
deviation from normality, presence of clusters, non-linear structure, or
other features of interest. For a fixed sample of data, PP finds the
orthonormal basis \(A\) that maximises the index value of the
projection, \(Y = XA\):

@@

Optimisation of PP is often discussed when new indexes are proposed
(Posse 1995; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
index, when they introduced the PPGT, which monitors the optimisation
visually so that the user can see the projected data leading in and out
of the optima. An implementation is available in the \texttt{tourr}
package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
(2021) illustrated how to diagnose optimisation processes, particularly
focusing on the guided tour, and revealed a need for improved
optimisation. While improving the quality of the optimisation solutions
in the tour is essential, it is also important to be able to view the
data projections as the optimisation progresses. Integrating the guided
tour with a global optimisation algorithm that is efficient in finding
the global optimal and enables viewing of the projected data during the
exploration process is a goal.

@@

\begin{itemize}
\tightlist
\item
  a pseudo-derivative, that searches locally for the best direction,
  based on differencing the index values for very close projections.
\item
  a brute-force optimisation (CRS).
\item
  a modified brute force algorithm described in Posse (1995).
\item
  an essentially simulated annealing (Bertsimas and Tsitsiklis 1993)
  where the search space is reduced as the optimisation.
\item
  a very localised search, to take tiny steps to get closer to the local
  maximum.
\end{itemize}

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, breakable, toprule=.15mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colback=white, bottomtitle=1mm, opacityback=0, rightrule=.15mm, toptitle=1mm, coltitle=black]

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abdulah2023large}{}}%
Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
Genton, and Ying Sun. 2023. {``Large-Scale Environmental Data Science
with {ExaGeoStatR}.''} \emph{Environmetrics} 34 (1): e2770.
\url{https://doi.org/10.1002/env.2770}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
Friedman, J. H., and J. W. Tukey. 1974. {``A Projection Pursuit
Algorithm for Exploratory Data Analysis.''} \emph{IEEE Transactions on
Computers} C-23 (9): 881--90.
\url{https://doi.org/10.1109/T-C.1974.224051}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-grochowski2011}{}}%
Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
Based on Quality of Projected Clusters.''} In \emph{Adaptive and Natural
Computing Algorithms}, edited by Andrej Dobnikar, Uroš Lotrič, and
Branko Šter, 89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-642-20267-4_10}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{The Annals of
Statistics} 13 (2): 435--75.
\url{https://doi.org/10.1214/aos/1176349519}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-karvonen2023asymptotic}{}}%
Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
Estimates in {G}aussian Process Interpolation.''} \emph{SIAM/ASA Journal
on Uncertainty Quantification} 11 (4): 1225--57.
\url{https://doi.org/10.1137/22M149288X}.

\leavevmode\vadjust pre{\hypertarget{ref-kr69}{}}%
Kruskal, J. B. 1969. {``Toward a Practical Method Which Helps Uncover
the Structure of a Set of Observations by Finding the Line
Transformation Which Optimizes a New {`Index of Condensation'}.''} In
\emph{Statistical Computation}, edited by R. C. Milton and J. A. Nelder,
427--40. New York: Academic Press.
\url{https://doi.org/10.1016/B978-0-12-498150-8.50024-0}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-lee2010projection}{}}%
Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
Large {\(p\)} Small {\(n\)} Data.''} \emph{Statistics and Computing} 20
(3): 381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2018}{}}%
Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
Computational Approach.''} \emph{Computational Statistics and Data
Analysis} 120 (C): 42--57.
\url{https://doi.org/10.1016/j.csda.2017.11.001}.

\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2020}{}}%
---------. 2020. {``Kurtosis-Based Projection Pursuit for Outlier
Detection in Financial Time Series.''} \emph{The European Journal of
Finance} 26 (2-3): 142--64.
\url{https://doi.org/10.1080/1351847X.2019.1647864}.

\leavevmode\vadjust pre{\hypertarget{ref-marie-sainte2010}{}}%
Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
{``An Efficient Optimization Method for Revealing Local Optima of
Projection Pursuit Indices.''} In \emph{Swarm Intelligence}, edited by
Marco Dorigo, Mauro Birattari, Gianni A. Di Caro, René Doursat, Andries
P Engelbrecht, Dario Floreano, Luca Maria Gambardella, et al., 60--71.
Berlin, Heidelberg: Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-642-15461-4_6}.

\leavevmode\vadjust pre{\hypertarget{ref-porcu2024matern}{}}%
Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
2024. {``The {M}atérn Model: A Journey Through Statistics, Numerical
Analysis and Machine Learning.''} \emph{Statistical Science} 39 (3):
469--92. \url{https://doi.org/10.1214/24-STS923}.

\leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
Posse, Christian. 1995. {``Projection Pursuit Exploratory Data
Analysis.''} \emph{Computational Statistics and Data Analysis} 20 (6):
669--87. \url{https://doi.org/10.1016/0167-9473(95)00002-8}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tourr}{}}%
Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
Package for Exploring Multivariate Data with Projections.''}
\emph{Journal of Statistical Software} 40 (2): 1--18.
\url{https://doi.org/10.18637/jss.v040.i02}.
